A Study on Common Bug Fix Patterns in Rust

### 2022-09-21 sections/methodology.tex
edits

The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints. We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms. Work in this vein makes the assumption that each cluster corresponds to one change pattern.

Our pipeline for finding pervasive patterns in Rust programs also follows this idea. We decided to implement our pipeline completely in Python. To make this work, we need a way to get an AST as a Python dictionary. Syn \footnote{\url{https://crates.io/crates/syn}} is a Rust crate which is designed to be employed in Rust procedural macros. However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.

% TODO Define change
A change pattern categorizes a class of changes. We thus need to define what a change is. For our purposes, a change is ....
To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it. After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries. A tree diff algorithm would specify the differences between two arbitrary trees. If the trees are abstracts syntax trees of programs, we simply call it an ASTDiff. We use dictdiff to obtain ASTDiffs.

An ASTDiff may include an arbitrary number of semantic changes. We are interested in finding the most important change within an ASTDiff and encoding that change in our datapoints. Sections \ref{sec:path_extraction} and \ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then helped us to obtain clusters that contained similar datapoints.

We used Pydriller, which is a repository miner for Python. Our target repositories were the top 18 most starred open source Rust projects on GitHub. We mined all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm on the obtained datapoints. In Section \ref{sec:clustering_data} we discuss why we chose DBSCAN for clustering and how we tuned it to improve clusters quality.

@@

As mentioned above, we use Syn for parsing Rust. Syn handles all of Rust. Syn's nonterminals are primarily divided and implemented through two groups: Structs and Enums, which can be found on Syn webpage. The union of the item described in these two groups constitutes the list of our non terminals (\verb+NT+). The dimensions of our datapoints have been partly designed in a way that they adhere to the elements of \verb+NT+. The other part of our dimenstions is comprised of BC-related lexemes (\verb+BCL+), e.g. clone, Rc, Box, etc. The full list of \verb+BCL+ elements can be found within our artifact. We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.

We parse two versions of a Rust changed file in a commit; the file before the commit and after it. We will obtain two trees created by Syn. The change that transforms the first tree to the second one yields the fix which was applied in the commit. To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries. Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the Rust AST into Python dictionaries. After obtaining the syntax trees in Python dictionary format, we can use dictdiff to determine the ASTDiff.

dictdiff tool provides modifications' types: 
\begin{itemize}
\item `add': A new structure has been added to the tree; 
\item `remove': A structure has been dropped from the tree; 
\item `change': The content of a sub structure of the first tree has changed.
\end{itemize}
dictdiff also determines the context within which the modification has occured. This context happens to be the root elements in our file, e.g. function blocks, struct blocks, impl blocks, etc. The modified content within a context is still in the form of tree. Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
% TODO rework the above paragraph

>>>>

% Let's start this by talking right away about changes rather than just ASTs
The general idea for automatically finding the pervasive code patterns within a list of programs is to parse the programs and encode the important parts of the derived AST in fixed-size datapoints. We need fixed-size datapoints, as it is a requirement for both centroid-based and density-based well-known clustering algorithms. Work in this vein makes the assumption that each cluster corresponds to one change pattern.

Our pipeline for finding pervasive patterns in Rust programs also follows this idea. We decided to implement our pipeline completely in Python. To make this work, we need a way to get an AST as a Python dictionary. Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate which aims to serve Rust procedural macros. However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries.

% TODO Define change
A change pattern categorizes a class of changes. We thus need to define what a change is. For our purposes, a change is ....
To compute the contents of a change, we need two code revisions: the revision before the commit, and the revision after it. After parsing these two revisions, we will have two ASTs, encoded in two Python dictionaries. Tree diff algorithms can compute the differences between two arbitrary trees. If the trees are programs' abstract syntax trees, we call their diff an ASTDiff. We use dictdiff to obtain ASTDiffs.

An ASTDiff may include an arbitrary number of semantic changes, although a best practice is to include only one semantic change in a commit. Because that best practice is not universally followed, we are interested in finding the most important change within an ASTDiff, and encoding that change in our datapoints. Sections~\ref{sec:path_extraction}~and~\ref{sec:weighting_scheme} provide a detailed explanation of how we select the most important semantic information, which then allowed us to obtain clusters that contained similar datapoints.

Our target repositories were the top 18 most-starred open source Rust projects on GitHub. Using Pydriller, a repository miner for Python, we mined all the bug related commits and ran them through our pipeline. Then, we applied the DBSCAN clustering algorithm on the resulting datapoints. In Section~\ref{sec:clustering_data} we justify our choice of DBSCAN for clustering, and how we tuned it to improve cluster quality.

@@

% TODO: BCL elements goes in an appendix.
As mentioned above, we use Syn for parsing Rust. Syn handles all of Rust. Each of our datapoints specifies whether any Syn non terminal (\texttt{NT}) is present or absent, i.e. containing one dimension per non terminal. However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related lexemes (BCL) that we identified, e.g. \texttt{clone}, \texttt{Rc}, \texttt{Box}. The full list of \verb+BCL+ elements can be found within our artifact. 
% I think we don't need this sentence here. Let's see what we do later.
%We define dimension items as \verb+DI = NT+ $\cup$ \verb+BCL+, as this set also equals the columns of our tables.

We parse two versions of a changed Rust file in a commit: the file before the commit and after it. This yields two Syn trees. Using the PLY tool (a Python implementation of lex and yacc), we wrote a simple transformer from the serialized Rust AST to Python dictionaries. The change that transforms the first tree to the second one is the fix that was applied in the commit. To find this difference, we use dictdiff, a Python library to find the diff of two Python dictionaries. Its output, in our context, is the ASTDiff.

dictdiff provides modifications' types: 
\begin{itemize}
\item `add': A new structure has been added to the tree; 
\item `remove': A structure has been dropped from the tree; 
\item `change': The content of a sub structure of the first tree has changed.
\end{itemize}
dictdiff also determines the context within which the modification has occured. This context happens to be the root elements in our file, e.g. function blocks, struct blocks, impl blocks, etc. The modified content within a context is still in the form of tree. Our objective right now is to find primary elements in our diff that describe the semantics of occurred change.
% TODO rework the above paragraph
